---
title: "Movie-Recommendation-System-R"
author: "Hoai Do"
date: "12/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### I. Importing Essential Libraries
Using the `recommenderlab`, `data.table`,`ggplot2`, and `reshape2` packages.
```{r, warning = FALSE, message = FALSE}
if(!require("recommenderlab")) install.packages("recommenderlab")
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```
Reading the `movies.csv` and `ratings.csv`. Deploying the `str()` function to display information about the dataset

```{r}
library(readr)
mv_data <- read.csv("movies.csv")
rt_data <- read.csv("ratings.csv")
str(mv_data)
```
```{r}
summary(mv_data)
```
```{r}
summary(rt_data)
```

### Data Pre-processing
I will first create a one-hot encoding to create a matrix that comprises of corresponding genres for each of the films.
```{r, warning = FALSE, message=FALSE}
movie_genre <- as.data.frame(mv_data$genres, stringsAsFactors=FALSE)
library(data.table)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE) #DataFlair
colnames(movie_genre2) <- c(1:10)
list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre
for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col]) #Author DataFlair
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)
```

In the next step of Data Pre-processing of R project, I will create a ‘search matrix’ that will allow me to perform an easy search of the films by specifying the genre present in our list.
```{r}
SearchMatrix <- cbind(mv_data[,1:2], genre_mat2[])
head(SearchMatrix) 
```

There are movies that have several genres, for example, Jumanji, which is an Adventure film also falls under the genres of Comedy, Fantasy, and Children. This applies to the majority of the films.

For this movie recommendation system to make sense of our ratings through recommenderlabs, I need to convert our matrix into a sparse matrix one. This new matrix is of the class ‘realRatingMatrix’. This is performed as follows:

```{r, warning=FALSE, message=FALSE}
ratingMatrix <- dcast(rt_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```

Now let take a look at some of the critical parameters providing us various options for developing this movie recommender
```{r, warning=FALSE, message=FALSE}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```
```{r, warning=FALSE, message=FALSE}
lapply(recommendation_model, "[[", "description")
```
```{r, warning=FALSE, message=FALSE}
recommendation_model$IBCF_realRatingMatrix$parameters
```

### Exploring Identical Data
Collaborative Filtering involves suggesting movies to the users that are based on collecting preferences from many other users. For example, if a user A likes to watch action films and so does user B, then the movies that the user B will watch in the future will be recommended to A and vice-versa. Therefore, recommending movies is **dependent on creating a relationship of similarity between the two users.** With the help of `recommenderlab`, we can compute **similarities using various operators like cosine, pearson as well as jaccard.**

```{r, warning=FALSE, message=FALSE}
similarity_mat <- similarity(ratingMatrix[1:7, ],
                               method = "cosine",
                               which = "users")
as.matrix(similarity_mat)
image(as.matrix(similarity_mat), main = "User's Similarities")
```

In the above matrix, each row and column represents a user. I have taken seven users and each cell in this matrix represents the similarity that is shared between the two users.

Now, we delineate the similarity that is shared between the films
```{r, warning=FALSE, message=FALSE}
movie_similarity <- similarity(ratingMatrix[, 1:7], 
                               method = "cosine",
                               which = "items")
as.matrix(movie_similarity)
image(as.matrix(movie_similarity), main = "Movies similarity")
```

Now I gonna to extract and store the most unique ratings into a table
```{r, warning=FALSE, message=FALSE}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values) # extracting unique ratings
Table_of_Ratings <- table(rating_values) # creating a count of movie ratings
Table_of_Ratings
```

### Most Viewed Movies Visualization
I will first count the number of views in a film and then organize them in a table that would group them in descending order.
```{r, warning = FALSE, message=FALSE}
library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(mv_data,
                                         mv_data$movieId == table_views[index,1])$title)
}
table_views[1:7,]
```

```{r, warning=FALSE, message=FALSE}
ggplot(table_views[1:7, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = '#c5bffb') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  ggtitle("Total Views of the Top Films")
```

### Heatmap of Movie Ratings
```{r, warning = FALSE, message=FALSE}
library(RColorBrewer)
image(ratingMatrix[1:25, 1:25], 
      axes = FALSE,
      main = "Heatmap of the first 20 rows and 5 columns")
```

### Performing Data Preparation
* Selecting insightful data
* Normalizing data
* Binarizing the data
For finding useful data in our dataset, we have set the threshold for the minimum number of users who have rated a film as 50. This is also same for minimum number of views that are per film. This way, we have filtered a list of watched films from least-watched ones.
```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                             colCounts(ratingMatrix) > 50]
movie_ratings
```

From the above output of ‘movie_ratings’, we observe that there are 420 users and 447 films as opposed to the previous 668 users and 10325 films. We can now delineate our matrix of relevant users as follows

```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")
```

Now, I will visualize the distribution of the average ratings per user
```{r, warning=FALSE, message=FALSE}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("#a6dbf8"), col=I("#ffa700")) +
  ggtitle("Distribution of the average rating per user")
```

### Normalizing the Data
In the case of some users, there can be high ratings or low ratings provided to all of the watched films. This will act as a bias while implementing our model. In order to remove this, we normalize our data. Normalization is a data preparation procedure to standardize the numerical values in a column to a common scale value. This is done in such a way that there is no distortion in the range of values. Normalization transforms the average value of the ratings column to 0. We then plot a heatmap that delineates our normalized ratings.
```{r, warning = FALSE, message = FALSE}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)

image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                          colCounts(normalized_ratings) > minimum_users],
main = "Normalized Ratings of the Top Users")
```

### Data Binarization
Binarizing the data means that we have two discrete values 1 and 0, which will allow our recommendation systems to work more efficiently. We will define a matrix that will consist of 1 if the rating is above 3 and otherwise it will be 0.
```{r, warning=FALSE, message=FALSE}
binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.95)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.95)
#movies_watched <- binarize(movie_ratings, minRating = 1)
good_rated_films <- binarize(movie_ratings, minRating = 3)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
colCounts(movie_ratings) > binary_minimum_users],
main = "Heatmap of the top users and movies")
```

### The Item Based Collaborative Filtering System
I will develop our very own Item Based Collaborative Filtering System. This type of collaborative filtering finds similarity in the items based on the people’s ratings of them. The algorithm first builds a similar-items table of the customers who have purchased them into a combination of similar items. This is then fed into the recommendation system.

The similarity between single products and related products can be determined with the following algorithm:

* For each Item i1 present in the product catalog, purchased by customer C.
* And, for each item i2 also purchased by the customer C.
* Create record that the customer purchased items i1 and i2.
* Calculate the similarity between i1 and i2.
I will build this filtering system by splitting the dataset into 80% training set and 20% test set.

```{r, warning =FALSE, message=FALSE}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```

### Developing the Recommendation System
At the stage of project, we will together explore the various parameters of the Item Based Collaborative Filter. These parameters are default in nature. In the first step, **k denotes the number of items for computing their similarities. Here, k is equal to 30.** Therefore, the algorithm will now identify the k most similar items and store their number. We use the `cosine` method which is the default one but you can also use pearson method.
```{r, warning=FALSE, message=FALSE}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters
```
```{r, warning=FALSE, message=FALSE}
recommen_model <- Recommender(data = training_data,
                          method = "IBCF",
                          parameter = list(k = 30))
recommen_model
class(recommen_model)
```
```{r, warning=FALSE, message=FALSE}
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
```
```{r, warning=FALSE, message=FALSE}
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
   main = "Heatmap of the first rows and columns")
```

In the next step, We will carry out the sum of rows and columns with the similarity of the objects above 0. I will visualize the sum of columns through a distribution
```{r, warning=FALSE, message=FALSE}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)
sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("#c2edda"), col=I("#ffa700"))+ ggtitle("Distribution of the column count")
```

We will create a top_recommendations variable which will be initialized to 10, specifying the number of films to each user. We will then use the predict() function that will identify similar items and will rank them appropriately. Here, each rating is used as a weight. Each weight is multiplied with related similarities. Finally, everything is added in the end.
```{r}
top_recommendations <- 10 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                          newdata = testing_data,
                          n = top_recommendations)
predicted_recommendations
```
```{r}
user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(mv_data,
                                         mv_data$movieId == movies_user1[index])$title)
}
movies_user2
```
```{r}
recommendation_matrix <- sapply(predicted_recommendations@items,
                      function(x){ as.integer(colnames(movie_ratings)[x]) }) # matrix with the recommendations for each user
recommendation_matrix[,1:4]
```
```{r, warning = FALSE, message=FALSE}
number_of_items <- factor(table(recommendation_matrix))
chart_title <- "Distribution of the Number of Items for IBCF"
qplot(number_of_items, fill=I("#c2edda"), col=I("#ffa700")) + ggtitle(chart_title)
```

Print out sample movie recommendation for 7 users with particular identity.
```{r, warning = FALSE, message=FALSE}
number_of_items_sorted <- sort(number_of_items, decreasing = TRUE)
number_of_items_top <- head(number_of_items_sorted, n = 7)
table_top <- data.frame(as.integer(names(number_of_items_top)),
number_of_items_top)
for(i in 1:7) {
table_top[i,1] <- as.character(subset(mv_data,
mv_data$movieId == table_top[i,1])$title)
}
colnames(table_top) <- c("Movie Title", "No. of Items")
head(table_top)
```